Human-level control through deep reinforcement learning

### 2017-12-27 article/article.tex
[Article] Add details on 'why a NN' and on weights update

\documentclass[letterpaper]{article}
\usepackage{natbib,alifexi}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{hyperref}

@@

We designed the agent in such a way that it selects actions according to its beliefs about future rewards, and its only objective is to maximize its total
cumulative reward. More formally, this cumulative calculation starts from present time $t$ and is computed as follows:
\[
  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k \; \big| \; s_t = s, a_t = a, \pi\right].
\]
where $\pi = P(a | s)$ is a behavior policy, $t$ is the current step, $r_t$ is the reward for step $t$, $a_t$ is the action selected at time $t$,
and $\gamma$ is the discount factor. Thus, we can describe the Q-value as the present value of all future rewards at current step $t$,
with respect to the environment state and the action selected at $t$.

Because of the very high dimensionality of input visual features, it is not reasonable to consider using a table for storing all Q-values.
Instead, one can design a learnable function that maps pixel values to a single row of this table, thus a vector of Q-values whose length is the size
of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article~\citep{Mnih2015}, since feedforward Neural Networks
are known to be universal function approximators (Theorem 2 in~\cite{hornik1991approximation}).
A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers~\citep{lecun1998gradient}. One characteristic of CNNs is the
presence of convolutional filters that maps pixel luminance to more abstract features. Each filter (or kernel) is locally connected to its output, which
allows the convolutional layer to capture the local information of the pixels, contrary to what dense (fully-connected) layers do. This procedure is greatly
inspired by the notion of receptive field introduced by Hubel and Wiesel~\citep{Hubel1962}.
Let $Q(s, a; \theta_i)$ be the Q-value approximation function represented by the neural network, parameterized by its weights $\theta$ at time step $i$.

@@

Every 4 steps of the Q-learning, the agent builds a minibatch from its memory (either one of the three mechanisms we just mentioned):
this minibatch is used by the DQN to process one step of stochastic gradient descent. In this way the values of Q are estimated iteratively and
refined over time. The deep neural network incorporates minimal prior knowledge about the environment: the only assumption lies in the number of
outputs in the last hidden layer, which has been set as the number of actions. In every Atari game, the action space is discrete and finite: this allows
us to determine the output shape of the neural network depending on the game itself. The number of actions made available by the environment
cannot be inferred by the DQN during training phase.

Also we compared DQN, as well as other state-of-the-art deep reinforcement learning architectures, to simpler reinforcement learning techniques (including
random action selection) using a normalized performance metric. The latter is formalized as below:
\[
    P = 100 \times (\text{score} - RAS) / (HS - RAS)
\]
where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by our
Space Invaders professional gamer \textbf{Stanislas Gueniffey}. By normalizing in this manner, the random action selection score RAS results in
a normalized performance of 0\% and the human score results in a normalized performance of 100\%.

@@

Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}

We developed our models with the support of the Arcade Learning Environment~\citep{bellemare13arcade}. This gives the opportunity to either
retrieve the environment state as a raw image (for deep reinforcement learning research) or as a state of the RAM (for classic reinforcement learning
applications). Because of the very high dimensionality of raw images, we applied some preprocessing steps on them. These steps are also aimed at
removing some artifacts produced by the Atari architecture. More specifically, the flickering effect implies that some game sprites only appear
in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
Then the sample size has been reduced by a factor of three by mapping RBG values to the pixel luminance, using the following formula:
\[
    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B
\]
After mapping the three color channels to a unique channel, the number of pixels has also been reduced. Each input image as been rescaled to a
$84 \times 84$ pixel grid.

@@

@@

\footnotesize
\bibliographystyle{apalike}
\bibliography{article}

>>>>

\documentclass[letterpaper]{article}
\usepackage{natbib,alifexi}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{commath}

\usepackage{hyperref}
\usepackage{flushend}

@@

We designed the agent in such a way that it selects actions according to its beliefs about future rewards, and its only objective is to maximize its total
cumulative reward. More formally, this cumulative calculation starts from present time $t$ and is computed as follows:
\begin{equation}
  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k \; \big| \; s_t = s, a_t = a; \pi\right].
\end{equation}
where $\pi = P(a | s)$ is a behavior policy, $t$ is the current step, $r_t$ is the reward for step $t$, $a_t$ is the action selected at time $t$,
and $\gamma$ is the discount factor. Thus, we can describe the Q-value as the present value of all future rewards at current step $t$,
with respect to the environment state and the action selected at $t$.

Because of the very high dimensionality of input visual features, it is not reasonable to consider using a table for storing all Q-values.
Instead, one can design a learnable function that maps pixel values to a single row of this table, thus a vector of Q-values whose length is the size
of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article~\citep{mnih2015human}, since feedforward Neural Networks
are known to be universal function approximators (Theorem 2 in~\cite{hornik1991approximation}).
A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers~\citep{lecun1998gradient}. One characteristic of CNNs is the
presence of convolutional filters that maps pixel luminance to more abstract features. Each filter (or kernel) is locally connected to its output, which
allows the convolutional layer to capture the local information of the pixels, contrary to what dense (fully-connected) layers do. This procedure is greatly
inspired by the notion of receptive field introduced by Hubel and Wiesel~\citep{Hubel1962}.
Let $Q(s, a; \theta_i)$ be the Q-value approximation function represented by the neural network, parameterized by its weights $\theta$ at time step $i$.

@@

Every 4 steps of the Q-learning, the agent builds a minibatch from its memory (either one of the three mechanisms we just mentioned):
this minibatch is used by the DQN to process one step of stochastic gradient descent. In this way the values of Q are estimated iteratively and
refined over time. The deep neural network incorporates minimal prior knowledge about the environment: the only assumption lies in the number of
outputs in the last hidden layer, which has been set as the number of actions. In every Atari game, the action space is discrete, finite and constant over time:
this allows us to determine the output shape of the neural network depending on the game itself. The number of actions made available by the environment
cannot be inferred by the DQN during training phase.

Also we compared DQN, as well as other state-of-the-art deep reinforcement learning architectures, to simpler reinforcement learning techniques (including
random action selection) using a normalized performance metric. The latter is formalized as below:
\begin{equation}
    P = 100 \times (\text{score} - RAS) / (HS - RAS),
\end{equation}
where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by our
Space Invaders professional gamer \textbf{Stanislas Gueniffey}. By normalizing in this manner, the random action selection score RAS results in
a normalized performance of 0\% and the human score results in a normalized performance of 100\%.

@@

Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}.

We developed our models with the support of the Arcade Learning Environment~\citep{bellemare13arcade}. This gives the opportunity to either
retrieve the environment state as a raw image (for deep reinforcement learning research) or as a state of the RAM (for classic reinforcement learning
applications). Because of the very high dimensionality of raw images, we applied some preprocessing steps on them. These steps are also aimed at
removing some artifacts produced by the Atari architecture. More specifically, the flickering effect implies that some game sprites only appear
in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
Then the sample size has been reduced by a factor of three by mapping RGB values to the pixel luminance, using the following formula:
\begin{equation}
    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B.
\end{equation}

After mapping the three color channels to a unique channel, the number of pixels has also been reduced. Each input image as been rescaled to a
$84 \times 84$ pixel grid.

As estimating the whole $Q^*$ function is practically not performable, $Q^*$ is estimated by a non-linear function depending on $\theta$ such that:
\begin{equation}
	\forall s \in S : \forall a \in A(s) : Q(s, a; \theta) \simeq Q^*(s, a).
\end{equation}

In this situation, the non-linear function is the neural network, and $\theta$ represents its weights.

The interest of using a neural network is to be able to update the weights such that at each time step $t$, the weights are $\theta_t$, and:
\begin{equation}
	\forall s \in S : \forall a \in A(s) : Q(s, a; \theta_t) \xrightarrow[t \to +\infty]{} Q^*(s, a).
\end{equation}

Therefore, the NN is trained by minimizing the Huber loss~\citep{huber1964robust} with $\delta=1$ of the Bellman equation (by a -- stochastic -- gradient descent).
The Huber loss $H_\delta$, parameterized by $\delta \in {\mathbb R^*}^+$, is a continuous function (also $C^1$ but not $C^2$) defined by the MSE
on a small neighbourhood of $0$, and by the absolute value outside this neighbourhood:
\begin{equation}
	H_\delta : \mathbb R \to \mathbb R : x \mapsto \begin{cases}\frac 12x^2                                &\text{ if } \abs x \leq \delta,\\
	                                                            \delta\left(\abs x - \frac 12\delta\right) &\text{ otherwise.}\end{cases}
\end{equation}

The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and has (\todo {probably}) been shown to increase
stability.~\todo{check if this has a proper reference or drop it}

\subsection{Neural Network architecture}

@@

\subsection{Update of the weights}

In order to reduce oscillation in $Q$ estimations, two separate neural networks have been used: the first one, denoted $Q$ and called the \textit{action-value}
network is the one used in action prediction, and the second one denotes $\hat Q$ and called the \textit{target action-value} network is used when
updating the weights. The parameters of $Q$ are denoted by $\theta$ as previously, and the parameters of $\hat Q$ are denoted by $\theta^-$.

Every 4 steps, at time $t$, a minibatch of 32 past experiences $\left\{(s_{t_j}, a_{t_j}, r_{t_j}, s_{t_j+1})\right\}_{j=1}^{32}$ is selected uniformly from the
memory in order to perform a Stochastic Gradient Descent (SGD), meaning that the average of the gradients of the loss function on each experience is used to update
the weights. As mentioned above, the loss function is a Huber loss function and is defined as:
\begin{equation}
	H_1\left(y_j - Q(s_{t_j}, a_{t_j}; \theta)\right),
\end{equation}
with $y_j$ set either to $r_j$ if $s_{t_j}$ is a final state (i.e. either the loss of a game or a forced end of episode) or to
$r_j + \gamma\max_{a' \in A(s_{t_j})}\hat Q(s_{j+1}, a'; \theta^-)$ otherwise; therefore using the target action-value network.

The $\hat Q$ target action-value network is then copied back from the action value network $Q$ every 10,000 steps.

@@

%\newpage
\footnotesize
\bibliographystyle{apalike}
\bibliography{article}
